---
title: "p8105_hw3_ga2612"
author: "Lupe Antonio"
date: "10/26/2023"
output: 
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Problem 1

```{r}
#loading dataset
library(p8105.datasets)
data("instacart")

instacart <- instacart %>% as_tibble()
```



How many aisles are there, and which aisles are the most items ordered from?
Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).


The instacart dataset has `r nrow(instacart)` observations and `r ncol(instacart)` variables/columns. Each row in this dataset represents a single product ordered through instacart. Variables of this dataset include order, product, and user IDs. Additionally, there are variables corresponding to the order in which product was added to cart, the time of order, days since prior order. There are also variables that describe the product, including product name (i.e. Yogurt, Milk, Water), the aisle (i.e. yogurt, fresh vegetables) and department (i.e. dairy eggs, produce) where it is located and the corresponding aisle and department IDs. 


There are a total of `r instacart %>% pull()

